---
title: "Week 10 Code—complete confounding in 2^k design"
author: "Lucy Liu"
date: "12/04/2018"
output: 
  html_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(knitr)
library(kableExtra)
library(tibble)
library(dplyr)
```

# (2^4, 2^2) design with complete confounding

## Given confounding effects, find layout.

* Equation method of finding the key block or confounded effects.

If a treatment effect $A^{l_1}B^{l_2}\ldots N^{l_n}$ is confounded in a replicate then any treatment combination $a^{x_1}b^{x_2}\ldots n^{x_n}$ will belong to the key block if
$$l_1x_1+l_2x_2+\ldots+l_nx_n = 0(mod 2)$$


**EG:** Suppose effects BCD, ABD are confouned, write down the layout of one replicate.

## Given the key block design, find the confounding effects.

**EG:** Give the key block has treatment effects (1), (bd), (abc) and (acd), find the confounding effects.

## Compute total effects and main effects
```{r, echo = FALSE}
design <- c("( 1 )", "(bd)", "(abc)", "(acd)",
            "( a )", "(abd)", "(bc)", "( cd )", 
            "( b )", "( d )", "(ac)", "(abcd)",
            "( c )", "(bcd)", "(ab)", "(ad)")

set.seed(1203)
rep1.data <- data.frame(Block = rep(1:4, each = 4), 
                        Rep = rep(-1, 16 ),
                        y = sample(-3:4, size = 16, replace = T) + c(0, 7, rep(0, 14)),
                        design = design)
set.seed(1204)
rep2.data <- data.frame(Block = rep(5:8, each = 4), 
                        Rep = rep(1, 16 ),
                        y = sample(-3:4, size = 16, replace = T) + c(0, 7, rep(0, 14)),
                        design = design)

treat.combi <- expand.grid(A = c(-1, 1), B = c(-1, 1), C = c(-1, 1), D = c(-1, 1))
treatment <- c("( 1 )", "( a )", "( b )", "(ab)", "( c )", "(ac)", "(bc)", "(abc)", 
               "( d )", "(ad)", "(bd)", "(abd)", "( cd )", "(acd)", "(bcd)", "(abcd)")

treat.combi$design <- treatment
rep1.join <- inner_join(rep1.data, treat.combi, by ="design")
rep2.join <- inner_join(rep2.data, treat.combi, by ="design")
data.all <- rbind(rep1.join, rep2.join)

display1.data <- cbind(rep1.join[1:4, 4:3], rep1.join[(1:4)+4, 4:3],
                      rep1.join[(1:4)+8, 4:3], rep1.join[(1:4)+12, 4:3])
display2.data <- cbind(rep2.join[1:4, 4:3], rep2.join[(1:4)+4, 4:3],
                      rep2.join[(1:4)+8, 4:3], rep2.join[(1:4)+12, 4:3])
display.all <- cbind(display1.data, display2.data)

col.name = character()
col.name[seq(1,16,by = 2)] <- paste0("block", 1:8)
col.name[seq(2,16,by = 2)] <- paste0("y", 1:8)

kable(display.all, col.names = col.name) %>%
        kable_styling("striped", full_width=FALSE) %>%
        add_header_above(c("Replicate I" = 8, "Replicate II" = 8)) 
```


```{r}
k = 4; r = 2; block_size = 4; replicate_size = 16
signtable <- data.frame(trt.comb = design)
signtable$A <- c(-1, -1, 1, 1, 1, 1, -1, -1, 
                 -1, -1, 1, 1, -1, -1, 1, 1)
signtable$B <- c(-1, 1, 1, -1, -1, 1, 1, -1,
                 1, -1, -1, 1, -1, 1, 1, -1)
signtable$C <- c(-1, -1, 1, 1, -1, -1, 1, 1,
                 -1, -1, 1, 1, 1, 1, -1, -1)
signtable$D <- c(-1, 1, -1, 1, -1, 1, -1, 1,
                 -1, 1, -1, 1, -1, 1, -1, 1)
signtable <- signtable%>%
        mutate(AB = A*B, AD = A*D, BD = B*D, BC = B*C, CD = C*D) %>%
        mutate(ABC = AB*C, ACD = A*C*D) %>%
        mutate(ABCD = ABC*D) %>% as.tibble()
kable(signtable) %>% kable_styling("striped", full_width=FALSE)
total.yeild_x = rep1.join$y + rep2.join$y
total.effect_X = apply(signtable[,-1], 2, function(x) x%*%total.yeild_x) # total effect
main.effect_X = round(total.effect_X/2^(k-1)/r, 4)  # main effect
tibble(effect.name = names(total.effect_X), total.effect_X = total.effect_X, main.effect_X = main.effect_X)
```
## compute ANOVA table manually

```{r}
head(data.all)
SS_X = total.effect_X^2/2^k/r 
grand.total = sum(data.all$y)
SS_total = (2^k*r-1) * var(data.all$y)
block_sum <- data.all %>%
        group_by(Block) %>%
        summarise(block_sum = sum(y))
SS_block = sum(block_sum$block_sum^2)/block_size - grand.total^2/(2^k*r)
replicate_sum <- data.all %>%
        group_by(Rep) %>%
        summarise(replicate_sum = sum(y))
SS_rep = sum(replicate_sum$replicate_sum^2)/replicate_size - grand.total^2/(2^k*r) 
SS_block.within.rep = SS_block - SS_rep
pvalue = pf(q = (SS_X/1)/(62.375/12), df1 = 1, df2 = 12)
ANOVA.manual <- data.frame(VarSource = c(names(total.effect_X), "SS_rep", "SS_block.in.rep", "error", "total"),
                           df = c(rep(1,12), r-1, (8-1) - (r-1), 32-13-1-6, 32), 
                           SS = c(SS_X, SS_rep, SS_block.within.rep, SS_total-sum(SS_X, SS_block),SS_total)) %>%
        mutate(MS = SS / df) %>%
        mutate(F_obs = MS/((SS_total-sum(SS_X, SS_block))/12))%>%
        mutate(Pvalue = pf(F_obs, df1 = 1, df2 = 12, lower.tail = F))
ANOVA.manual$F_obs[15:16] = NA
ANOVA.manual$Pvalue[15:16] = NA
print(ANOVA.manual, digits = 4)
```
## compute ANOVA table in R

```{r}
lm.fit <- lm(y ~ A*B*C*D - B:C:D - A:B:D - A:C + Rep + factor(Block), data = data.all)
(ANOVA.table <- anova(lm.fit))
```

## Suitable model

According to the Pvalues in the above ANOVA table, only main effects, A and C, are significant, we reduce the model to be

$$y = \mu + A +  C + \epsilon$$
