---
title: "Week 5 Code--General two-way ANOVA model"
author: "Lucy Liu"
date: "10/29/2018"
output: 
  html_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(knitr)
library(kableExtra)
library(tibble)
library(dplyr)
reaction.df <- read.table("reaction.time.txt", header = T) %>% 
        as.tibble() %>%
        mutate_at(.vars = vars(Trtmt:B), .funs = as.factor) %>%
        select(-Order, -Trtmt) 
names(reaction.df) <- c("Simulus", "ElapsedTime", "ReactionTime")
reaction.df
```

# General two-way ANOVA model

$$Y_{ijk} = \mu + \alpha_i + \beta_j + (\alpha\beta)_{ij} + \epsilon_{ijk},\quad i = 1,2,\quad j = 1,2,3,\quad k = 1,2,3$$
where $\epsilon_{ijk}\sim N(0, \sigma^2).$

**Parameters in the model:**

* $\mu$ = overall mean
* $\alpha_i$ = main effect of ith cue stimulus
* $\beta_j$ = main effect of jth  elapsed time
* $(\alpha\beta)_{ij}$ = interaction effect due to combination (i,j) of the two factors
* $\sigma^2$ = variance of random error

## Estimation of Parameters 

```{r}
options(contrasts=c("contr.sum","contr.poly"))
lm.fit <- lm(ReactionTime ~ Simulus*ElapsedTime, data = reaction.df)
summary(lm.fit)
```

## ANOVA table

```{r}
aov.fit <- aov(ReactionTime ~ Simulus*ElapsedTime, data = reaction.df)
summary(aov.fit)
```

**Interpretation: **

Test for interaction effects of Stimulus and ElapsedTime:

* $H_{0AB}: (\alpha\beta)_{ij} = 0$ for all i,j
* $F_{obs} = 1.463$
* Pvalue =  0.270 > 0.05
* Fail to reject $H_{0AB}$ => reduce the model to only include two main effects => continue your analysis in LabA Q3(g).

# Compute by hand

Use the following quantities, continue to calculate the estimate of parameters, SS and MSE. Solutions  will be shown on board.

```{r}
(mu.hat <- mean(reaction.df$ReactionTime)) # overall mean y_{...}
(y_i.. <- reaction.df %>%  # y_{i..}
        group_by(Simulus) %>%
        summarise(ReactionTime_mean = mean(ReactionTime))) 
(y_.j. <- reaction.df %>%
        group_by(ElapsedTime) %>%
        summarise(ReactionTime_mean = mean(ReactionTime))) # y_{.j.}
y_ij._var <- reaction.df %>%
        group_by(Simulus, ElapsedTime) %>%
        summarise(ReactionTime_mean = mean(ReactionTime), ReactionTime_var = var(ReactionTime))
mean_varTB <- data.frame(df1 = y_ij._var[1:3, 3:4], df2 = y_ij._var[4:6, 3:4]) 
rownames(mean_varTB) = c("ElapsedTime1","ElapsedTime2","ElapsedTime3")

options(knitr.table.format = "html") 
kable(mean_varTB, col.names = rep(c("mean", "variance"),2), digits = 8) %>%
        kable_styling(full_width = F, position = "left") %>%
        kable_styling("striped") %>%
        add_header_above(c(" " = 1, "Simulus 1" = 2, "Simulus 2" = 2))
```




## Estimation of parameters


## ANOVA table
```{r}
cat("Also given SS_{total} is ",  var(reaction.df$ReactionTime)*(nrow(reaction.df)-1))
```

## Draw interaction plots

## Contrasts ( Bonferroni correction method)

Assume the interaction term is significant and we are interested in `Simulus` effect. Therefore, we need to construct CI at three levels of `ElapsedTime`.

Will be shown on board.
```{r}
qt(p = 1 - 0.05/2/3, df = 12) # W_{B}
MSE <- sum(aov.fit$residuals^2)/12
sqrt(2 * MSE/3)
```

# Q&A in LabA