---
title: "Week 2 Code"
author: "Lucy Liu"
date: "10/9/2018"
output: 
  html_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

# ANOVA table and test effects (calculation by hand)

This will be showed on board in section.

# Analysis CRD in R

## ANOVA table and test effects (calculation in R)

* The objective of the analysis is to check if all the $\nu$ treatments have the same effect, that is we want to test
$$H_0: \tau_1= \tau_2= \ldots =\tau_\nu$$

```{r}
battery.data <- read.table ("battery.txt" , header = TRUE) 
battery.data$fType <- factor(battery.data$Type)
model1 <- aov(LPUC ~ fType, data = battery.data) # Fit aov model anova(model1) 
# ANOVA table and F test
summary(model1)
```

##  Single Contrasts

* suppose we are interested in the following contrasts:
$$\tau_1-\tau_2$$
$$\frac{1}{2}(\tau_1+\tau_2) -\frac{1}{2}(\tau_3+\tau_4)$$
$$\frac{1}{2}(\tau_1+\tau_2) - \tau_4$$
```{r}
# install.packages("contrast")
library(contrast)
fit1 <- lm(LPUC ~ fType, data = battery.data) # Fit aov model 
contrast(fit1, list(fType=as.factor(1)), list(fType=as.factor(2)))
contrast(fit1, list(fType=as.factor(1:2)), list(fType=as.factor(3:4)), type="average")
contrast(fit1, list(fType=as.factor(1:2)), list(fType=as.factor(4)), type="average")
```
## Multiple comparison problem

* the experimenter does m preplanned simultaneous comparisons.

### Bonferroni correction method for m multiple comparisons
$$\sum c_i\tau_i \in(\sum c_i\bar{y}_{i.} -w_B\sqrt{MSE\sum c_i^2/r_i}, \sum c_i\bar{y}_{i.} +w_B\sqrt{MSE\sum c_i^2/r_i})$$
where $W_B=t_{n-\nu,\alpha/2m}$.

 Suppose m = 5, one of the 5 comparisons aims to estimate:
 $$\frac{1}{2}(\tau_1+\tau_2) -\frac{1}{2}(\tau_3+\tau_4)$$
 Therefore, $c_1=c_2=\frac{1}{2}, c_3=c_4=-\frac{1}{2}$.
```{r}
# install.packages("plyr")
w.B <- qt(p = 1-(0.05/2/5), df = 16-4)
library(plyr)
(mean.table <- ddply(battery.data, .(fType), summarize, mean = mean(LPUC)))
c.vector <- c(0.5, 0.5, -0.5, -0.5)
# the Bonferroni confidence interval is 
c(mean.table$mean%*%c.vector-w.B*sqrt(2368*sum(c.vector^2)/4), mean.table$mean%*%c.vector+w.B*sqrt(2368*sum(c.vector^2)/4))
```

### Scheffe correction method for all multiple comparisons
$w_S = \sqrt{(\nu-1)*F_{\nu-1, n-\nu,\alpha}}$
```{r}
# the Scheffe confidence interval is 
w.S <- sqrt((4-1) * qf(p = 0.95, df1 = 4-1, df2 = 16-4))
c(mean.table$mean%*%c.vector-w.S*sqrt(2368*sum(c.vector^2)/4), mean.table$mean%*%c.vector+w.S*sqrt(2368*sum(c.vector^2)/4))
```
### Tukey method for all pairwise comparisons
```{r}
TukeyHSD(model1, "fType", ordered = TRUE, conf.level = 0.95)
plot(TukeyHSD(model1, "fType", ordered = TRUE, conf.level = 0.95))
```

