---
title: "Week4 Code"
author: "Lucy Liu"
date: "10/24/2017"
output: 
  html_document:
    number_sections: yes
    toc: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE,fig.align = "center"
                      ,fig.height=10,fig.width=14)
```



# Residual Diagnostic plots 
```{r}
#################                 Simulated models              #################
set.seed(20171025)
x<-rnorm(n = 100, sd = 1, mean = seq(1,10,length.out = 100)) # predictor x
### linear model statisfies four assumptions
y.perfect<-1+2*x+rnorm(n=100,mean = 0,sd = 1)
y.perfect_model<-lm(y.perfect~x)
### quadratic model, i.e. violates linear assumption
y.quadratic<-1+2*x+3*x^2+rnorm(n=100,mean = 0,sd = 1)
y.quadratic_model<-lm(y.quadratic~x)
### linear model with Var(epsilon_i)=f(x_i), i.e. violates constant assumption
y.nonconstant.var<-1+2*x+rnorm(n=100,mean = 0,sd = x)
y.nonconstant.var_model<-lm(y.nonconstant.var~x)
### linear model with epsilon_i is t(2), i.e. violates normal assumption
y.nonnormal.var<-1+2*x+rt(n = 100,df = 2)
y.nonnormal.var_model<-lm(y.nonnormal.var~x)
#################               scatter plot of y~x             #################
par(mfrow=c(2,2))
title<-c(expression(paste(y==1+2*x+epsilon, ", ",epsilon,"~" ,N(0,x^2)) ),
         expression(paste(y==1+2*x+epsilon, ", ",epsilon,"~" ,"t(2)") ),
         expression(paste(y==1+2*x+epsilon, ", ",epsilon,"~" ,"N(0,1)") ),
         expression(paste(y==1+2*x+3*x^2+epsilon, ", ",epsilon,"~" ,"N(0,1)")))
color<-c(3,4,1,2)
plot(x, y.nonconstant.var, pch=16, col=color[1],main=title[1])
plot(x, y.nonnormal.var,pch=16,col=color[2],main=title[2])
plot(x, y.perfect, pch=16, col=color[3],main=title[3])
plot(x, y.quadratic, pch=16, col=color[4],main=title[4])
```

## Residual vs. fitted
* Check assumptions: $E(\varepsilon_i)=0 \text{ and } Var(\varepsilon_i)=\sigma^2$
* Reasonable $e_i$ vs. $\hat{y_i}$ graph: should have mean zero and constant range.  
We don't expect the variability of the residuals to change over the range of the dependent variable, so there shouldn't be any discernable pattern to this plot. 
* Plots of the simulated cases are as followings.
```{r}
modelname.list<-grep("_model",x =ls(),value = T )
model.list<-list()
for (i in 1:length(modelname.list)){
      model.list[[i]]<-get(modelname.list[i])
}
par(mfrow=c(2,2))
for( i in 1:4) plot(model.list[[i]],which=1,col=color[i],caption =title[i])
```

## QQ plots
* Check normality: $\varepsilon_i\sim N(0,\sigma^2)$
* Reasonable QQ-plot: should follow a straight line. 
* Plots of the simulated cases are as followings.
```{r}
par(mfrow=c(2,2))
for (i in 1:4) plot(model.list[[i]],which = 2,col=color[i],main =title[i])
```

## Scale-location plots
* Check assumptions: $E(\varepsilon_i)=0 \text{ and } Var(\varepsilon_i)=\sigma^2$
* Reasonable $\sqrt{r_i}$ vs. $\hat{y_i}$ graph: should have mean zero and constant range, which is similar to the residuals versus fitted values plot.
* Plots of the simulated cases are as followings.
```{r}
par(mfrow=c(2,2))
for (i in 1:4) plot(model.list[[i]],which = 3,col=color[i],main =title[i])
```

# Box Cox transformation 
```{r,fig.height=5,fig.width=8}
### simulated model log(y)= 1+0.02*x+\epsilon
set.seed(20171025)
y.log<-exp(0.1+0.02*x+rnorm(100,sd=1))
logY.model<-lm(y.log~x)
library(car)
logY.boxcox<-boxCox(logY.model)
(lambda.opt = logY.boxcox$x[which.max(logY.boxcox$y)])
```
Though the optimal $\lambda$ is -0.06, the maximum likelihood estimate should be rounded to an interpretable value. Also, the  $\lambda$ value you choose should be in the confidence interval, i.e. within the dotted lines. For this example,  the best lambda is close to 0 and 0 is within the confidence interval, then choose the logarithm which is corresponds to when $\lambda$ is 0 in Boxcox transformation.

# Exploratory analysis for  multiple linear regression

## Scatterplot Matrix
The scatterplot matrix can provide answers to the following questions:

1. Are there pairwise relationships between the variables?
2. If there are relationships, what is the nature of these relationships?
3. Are there outliers in the data?
```{r}
library(faraway)
colnames(teengamb)
pairs(~status+income+verbal+gamble,data = teengamb) # do NOT include categorical variables!
```

## Added variable plots
 
```{r}
mul.model<-lm(gamble ~.,data = teengamb) # gamble is the expenditure on gambling in pounds per year
avPlots(mul.model)
summary(mul.model) # relate slopes of added variable plots with coefficients in the fitted model
```

**Interpretation of added variable plots:**

For example, added variable plot for `income` (in pounds per week):

* Regression `gamble` on all $x$ except `income` and get residuals $\hat{\delta}$. These represent `gamble` with the other $X$-effects taken out.
* Regression `income` on all $x$ except `income` and get residuals $\hat{\gamma}$. These represent `income` with the other $X$-effects taken out.
* The `added variable plot for income` shows $\hat{\delta}$ vs. $\hat{\delta}$.
* Aim to look for nonlinear pattern, outliers.


