---
title: "Week3 Code"
author: "Lucy Liu"
date: "10/17/2017"
output: 
  html_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE,fig.align = "center")
```

# Questions in Homework 2 
```{r}
library(faraway)
linear.model<-lm(gamble ~ income, data = teengamb) # simple regression case
summary(linear.model)
```
## 2(c) Information in Summary 

* estimate of $\sigma^2$: $\hat{\sigma^2}=24.95^2$.
* estimated standard errors of $\beta_0$ and $\beta_1$: 6.030 and 1.036.
* t-test for $H_0: \beta_0=0$, $t_{obs}=-1.049$, P-value = 0.3.
* t-test for $H_0: \beta_1=0$, $t_{obs}= 5.330$, P-value = 3.05e-06.
```{r}
vcov(linear.model)
```

* estimated covariance between these two estimates $\hat{\beta}_0$ and $\hat{\beta}_1$ is -4.98.

## 3(c) $R^2$ explanation

* $R^2$ coefficient is 0.387, which is also called the coefficient of determination.  
Explanation: 38.7\% of the variation in gamble expenditure is 'explained by' the variation in predictor income.

## 4(c) 99% prediction interval
```{r}
pointwise.obs<-predict(linear.model,newdata = data.frame(income=8),interval = "prediction",level = 0.99,se=T)
pointwise.obs$fit
```

Then predicted value is 37.84, and 99% prediction interval is  (-30.61333, 106.292).

# Simultaneous confidence intervals for mean response 

## Bonferroni correction
If we form $k$ confidence intervals with overall confidence level $100(1-\alpha)\%$, formulate each individual CI with level $100(1-\alpha/k)\%$.
$$\hat{Y}(x_0)\pm t_{n-2, 1-\frac{\alpha}{2k}}SE(\hat{Y}(x_0))$$.
```{r}
CI.bonferroni<-predict(linear.model,newdata = data.frame(income=c(2,3,4,5)),level = 1-0.05/4,interval = "confidence",se=T)
print(CI.bonferroni$fit[,2:3]) # Bonferroni CI, significence level for each CI is 0.05/4
```


## Scheffe's method

Replace $t_{n-2,1-\frac{\alpha}{2}}$ with $\sqrt{2F_{2,n-2, 1-\alpha}}$, then
$$\hat{Y}(x_0)\pm \sqrt{2F_{2,n-2, 1-\alpha}} SE(\hat{Y}(x_0))$$ is the $100(1-\alpha)\%$ simultaneous confidence band. It holds for $\forall x_0\in R$.
```{r}
n=nrow(teengamb)
CI.normal<-predict(linear.model,newdata = data.frame(income=c(2,3,4,5)),level = 1-0.05,interval = "confidence",se=T)
data.frame(lower.bound=c(CI.normal$fit[,1]-sqrt(2*qf(1-0.05,2,n-2))*CI.normal$se.fit),
           upper.bound=c(CI.normal$fit[,1]+sqrt(2*qf(1-0.05,2,n-2))*CI.normal$se.fit))
```


# ANOVA for regression
```{r}
anova(linear.model)
```
* F statistic =  28.407 for testing: $H_0:\beta_1=0$.  

***Note, in simple linear regression, F statistic =  28.407=t $\text{statistic}^2$ = $5.330^2$. And you get the same P-value for F test and t test, which are equivalent tests. But this is true only for simple linear regression.***

* P-value =3.045e-06.
