---
title: "Week2 Code"
author: "Lucy Liu"
date: "10/8/2017"
output: 
  html_document:
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE,fig.align = "center")
```

# R Basics---2 lm function
We use the dataset, `teengamb`, in `faraway` library to study teenage gambling in Britain. Specifically, we are going to explore the linear relationship between the teenagers' expenditure on gambling per year and income per week. 

## `lm` function and its output
```{r}
library(faraway)
linear.model<-lm(gamble ~ income, data = teengamb) # simple regression case
## lm(gamble ~ income-1,data = teengamb) # remove the slope
## lm(gamble ~., data = teengamb) # use all varibles except gamble as predictors
## lm(gamble ~ income+sex,data = teengamb) # use income and sex as predictors 
summary(linear.model)
```
**Output:**

*  standard errors: $\hat{\sigma}=24.95$
*  t-statistics and p-values: 

| | $\beta_0$|$\beta_1$|
| :---------------| :---------------|:-----------------|
t-statistic|-1.049| 5.330|
p-value| 0.3  |3.05$\times10^{-6}$
* $R^2$ coefficient is 0.387, which is also called the coefficient of determination.  
Explanation: 38.7\% of the variation in gamble expenditure is 'explained by' the variation in predictor income.

## Confidence intervals for coefficients.
```{r}
confi95<-confint(linear.model,level = 0.95)
confi95
```

## Prediction for new observations and mean response
```{r,fig.align="center",fig.width=10}
#######                  Prediction for a specific income value             #######
# pointwise.mean<-predict(linear.model,newdata = data.frame(income=8),interval = "confidence",se=T)
grid<-seq(min(teengamb$income),max(teengamb$income),length.out = 100)
pointwise.mean<-predict(linear.model,newdata = data.frame(income=grid),interval = "confidence",se.fit =T)
pointwise.obsv<-predict(linear.model,newdata = data.frame(income=grid),interval = "prediction",se.fit =T)
par(mfrow=c(1,2))
matplot(grid,pointwise.mean$fit,lty=c(1,2,2),col=c(1,2,2),type="l",
        xlab="income(pounds)",ylab="gamble(pounds)",lwd = 2,main="prediction for mean response")
matplot(grid,pointwise.obsv$fit,lty=c(1,2,2),col=c(1,2,2),type="l",
        xlab="income(pounds)",ylab="gamble(pounds)",lwd = 2,main="prediction for new observation")
```